---
title: "EBP Skills Assessment"
author: "Richard Campo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries}
library(tidyverse)
library(tidycensus)
library(sf)
library(tmap)
options(tigris_use_cache = TRUE) # cache census shapefiles
```

# Problem 1: Map average commute time and median HH income for census tracts in a metro area using ACS

First, I look for variables related to commute time in the ACS
```{r load-acs}
acs20 <- load_variables(2020, "acs5", cache = TRUE)
view(acs20)
```

I found B08303_001: Travel Time to Work, and found the definition in the Census's code book [here](https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2020_ACSSubjectDefinitions.pdf). The units are minutes per week.

I'll choose Cook County, IL since that's where I live.
```{r}
cook_commute <- get_acs(geography = "tract", 
                        variables = c(commute_time = "B08303_001"),
                        state = "Illinois",
                        county = "Cook", 
                        year = 2020,
                        geometry = TRUE)
```

Next, I found B19013_001: Median Household Income in the Past 12 months, so I'll pull that data.
```{r}
cook_med_inc <- get_acs(geography = "tract", 
                        variables = c(med_income = "B19013_001"),
                        state = "Illinois",
                        county = "Cook", 
                        year = 2020,
                        geometry = TRUE)
```

I tried plotting the commute data, but got an error that there's an empty geometry I need to remove first. I found [this post](https://gis.stackexchange.com/a/447255) on how to check for and remove empty geometries.
```{r}
# Filter out empty geometries
cook_commute <- cook_commute[!st_is_empty(cook_commute), ]

cook_med_inc <- cook_med_inc[!st_is_empty(cook_med_inc), ]
```

```{r}
commute_plot <- tm_shape(cook_commute) + 
  tm_fill("estimate", n = 4, palette = "Oranges", style = "jenks", 
          title = "Hours per Week") + 
  tm_borders(alpha = 0.4) + 
  tm_layout(legend.text.size = 0.6, legend.title.size = 1.0,
            legend.position = c("left", "center"), frame = FALSE,
            main.title = "Average Commute Time in Cook County 2020",
            main.title.size = 0.85) +
  tm_scale_bar(position = c("left", "bottom"))

med_inc_plot <- tm_shape(cook_med_inc) + 
  tm_fill("estimate", n = 4, palette = "Greens", style = "jenks", 
          title = "2020 Inflation-adjusted dollars") + 
  tm_borders(alpha = 0.4) + 
  tm_layout(legend.text.size = 0.6, legend.title.size = 0.9,
            legend.position = c("left", "center"), frame = FALSE,
            main.title = "Median Household Income in Cook County 2020",
            main.title.size = 0.85) +
  tm_scale_bar(position = c("left", "bottom"))

tmap_arrange(commute_plot, med_inc_plot, ncol = 2)
```

```{r}

```


```{r}


```

